### 2024-01-09 code/codeprocessmale.Rmd

---
title: "male code"
output: html_document
---

```{r packages}
# Load packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(cowplot)
```
```{r Functions}
# Read the CSV file
data <- read.csv("../gym2024data/cleandata/data_2022_2023.csv", header = TRUE)

#Convert all last names to lower case
data$LastName <- tolower(data$LastName)

#Remove rows with scores equal to zero
data <- data %>% filter(Score != 0)

#Combine VT, VT1, and VT2 in the "Apparatus" column
data$Apparatus <- ifelse(data$Apparatus %in% c("VT", "VT1", "VT2"), "VT", data$Apparatus)

#Calculate the mean and standard deviation of each male gymnast's score for each apparatus
male_scores <- data %>%
  filter(Gender == "m") %>%
  group_by(Apparatus, LastName, FirstName) %>%
  summarize(
    Mean_Score = mean(Score),
    SD_Score = sd(Score)
  )

#Calculate the mean and standard deviation of each United States male gymnast's score for each apparatus
usa_male_scores <- data %>%
  filter(Country == "USA" & Gender == "m") %>%
  group_by(Apparatus, LastName) %>%
  summarize(
    Mean_Score = mean(Score),
    SD_Score = sd(Score)
  )

#Add a "Number of Observations" column to the usa_male_scores table
usa_male_scores <- data %>%
  filter(Country == "USA" & Gender == "m") %>%
  group_by(Apparatus, LastName) %>%
  summarize(
    Num_Observations = n(),
    Mean_Score = mean(Score),
    SD_Score = sd(Score)
  )

#Create a table with the top 10 athletes with the highest mean scores for each apparatus and to plot each apparatus on separate graphs
  # Filter for USA male gymnasts
  usa_female_scores <- data %>%
  filter(Country == "USA" & Gender == "m")

  # Group by apparatus, calculate mean, standard deviation, and number of observations
  usa_male_scores <- usa_male_scores %>%
  group_by(Apparatus, LastName) %>%
  summarize(
    Num_Observations = n(),
    Mean_Score = mean(Score),
    SD_Score = sd(Score)
   )

  # Create a table with the top 10 athletes for each apparatus
  top_10_athletes <- usa_male_scores %>%
  group_by(Apparatus) %>%
  arrange(Apparatus, desc(Mean_Score)) %>%
  slice_head(n = 10)

  # Plot each apparatus on one graph
  for (apparatus in unique(top_10_athletes$Apparatus)) {
  data_subset <- filter(top_10_athletes, Apparatus == apparatus)
  p <- ggplot(top_10_athletes, aes(x = SD_Score, y = Mean_Score, label = LastName)) +
  geom_point() +
  geom_text(hjust = 0, vjust = 0) +
  labs(x = "Standard Deviation", y = "Mean Score") +
  facet_wrap(~Apparatus, scales = "free") +
  theme_minimal()
  print(p)
  }

  # Print the table of top 10 athletes
  top_10_athletes

#Add a "Mean minus Standard Deviation" column to the top_10_athletes table
  # Calculate the mean minus standard deviation
  top_10_athletes <- top_10_athletes %>%
  mutate(Mean_minus_SD = Mean_Score - SD_Score)

  # Print the table of top 10 athletes
  top_10_athletes

#Add a "Mean minus (0.5) x Standard Deviation" column to the top_10_athletes table
  # Calculate the mean minus (0.5) x standard deviation
top_10_athletes <- top_10_athletes %>%
  mutate(Mean_minus_0.5xSD = Mean_Score - 0.5 * SD_Score)

  # Print the table of top 10 athletes
  top_10_athletes

#Calculate the means and standard deviations for the specified US male athletes across all apparatuses
 # Filter for USA male athletes
  usa_male_scores <- data %>%
  filter(Country == "USA" & Gender == "m")

  # Define the list of last names
  target_last_names <- c("juda", "young", "moldauer", "hong", "hale", "malone", "whittenburg", "loos", "bock", "walker")

  # Filter the data for the specified athletes
  selected_athletes <- usa_male_scores %>%
  filter(LastName %in% target_last_names)

  # Group by athlete, apparatus, and calculate mean and standard deviation
  athlete_apparatus_scores <- selected_athletes %>%
  group_by(LastName, Apparatus) %>%
  summarize(
    Mean_Score = mean(Score),
    SD_Score = sd(Score)
  )

  # Print the results
  print(athlete_apparatus_scores)

#Calculate the sum of means and standard deviations for the targeted athletes from the athlete_apparatus_scores table
  # Define the list of last names
  target_last_names <- c("juda", "young", "moldauer", "hong", "hale", "malone", "whittenburg", "loos", "bock", "walker")

  # Filter the data for the specified athletes
  selected_athletes_scores <- athlete_apparatus_scores %>%
  filter(LastName %in% target_last_names)

  # Group by athlete and calculate the sums of means and standard deviations
  sums <- selected_athletes_scores %>%
  group_by(LastName) %>%
  summarize(
    Sum_Means = sum(Mean_Score),
    Sum_SD = sum(SD_Score)
  )

  # Print the sums for each athlete
  print(sums)

#Create a scatter plot of the summed means on the y-axis and summed standard deviations on the x-axis
  # Create a scatter plot
  p1 <- ggplot(sums, aes(x = Sum_SD, y = Sum_Means, label = LastName)) +
  geom_point() +
  geom_text(hjust = 0, vjust = 0) +
  labs(x = "Sum of Standard Deviations", y = "Sum of Means", title = "Male All-Around Candidates")

  # Display the plot
  print(p1)

#Calculate and display the mean score and standard deviation sum for every combination of 3 athletes from the sums table
  # Calculate all combinations of 3 athletes
  athlete_combinations <- combn(sums$LastName, 3)

  # Create a data frame to store the results
  combinations_results <- data.frame(
  Combination = character(choose(length(sums$LastName), 3)),
  Mean_Score = numeric(choose(length(sums$LastName), 3)),
  SD_Sum = numeric(choose(length(sums$LastName), 3))
  )

  # Calculate and store the mean and standard deviation sum for each combination
  for (i in 1:ncol(athlete_combinations)) {
  combination <- athlete_combinations[, i]
  mean_scores <- sums$Sum_Means[sums$LastName %in% combination]
  sd_sums <- sums$Sum_SD[sums$LastName %in% combination]
  combinations_results$Combination[i] <- paste(combination, collapse = ", ")
  combinations_results$Mean_Score[i] <- sum(mean_scores)
  combinations_results$SD_Sum[i] <- sum(sd_sums)
  }

  # Print the results
  print(combinations_results)

#Plot the mean (y-axis) vs. standard deviation (x-axis) for the 10 combinations with the highest mean score from the combinations_results table
  # Sort the combinations by mean score in descending order and select the top 10
  top_10_combinations <- head(combinations_results[order(-combinations_results$Mean_Score), ], 10)

  # Create a scatter plot for the top 10 combinations
  p2 <- ggplot(top_10_combinations, aes(x = SD_Sum, y = Mean_Score)) +
  geom_point() +
  geom_text(aes(label = Combination), hjust = 0, vjust = 0) +
  labs(x = "Sum of Standard Deviations", y = "Mean Score") +
  ggtitle("Top 10 Combinations of 3 Male Athletes with Highest Mean Score")

  # Display the plot
  print(p2)

#Calculate and display the mean score and standard deviation sum for every combination of 4 athletes from the sums table
  # Calculate all combinations of 4 athletes
  athlete_combinations_4 <- combn(sums$LastName, 4)

  # Create a data frame to store the results
  combinations_results_4 <- data.frame(
  Combination = character(choose(length(sums$LastName), 4)),
  Mean_Score = numeric(choose(length(sums$LastName), 4)),
  SD_Sum = numeric(choose(length(sums$LastName), 4))
  )

  # Calculate and store the mean and standard deviation sum for each combination
  for (i in 1:ncol(athlete_combinations_4)) {
  combination <- athlete_combinations_4[, i]
  mean_scores <- sums$Sum_Means[sums$LastName %in% combination]
  sd_sums <- sums$Sum_SD[sums$LastName %in% combination]
  combinations_results_4$Combination[i] <- paste(combination, collapse = ", ")
  combinations_results_4$Mean_Score[i] <- sum(mean_scores)
  combinations_results_4$SD_Sum[i] <- sum(sd_sums)
  }

  # Print the results
  print(combinations_results_4)

#Plot the mean (y-axis) vs. standard deviation (x-axis) for the 10 combinations with the highest mean score from the combinations_results_4 table
  # Sort the combinations by mean score in descending order and select the top 10
  top_10_combinations_4 <- head(combinations_results_4[order(-combinations_results_4$Mean_Score), ], 10)

  # Create a scatter plot for the top 10 combinations
  p3 <- ggplot(top_10_combinations_4, aes(x = SD_Sum, y = Mean_Score)) +
  geom_point() +
  geom_text(aes(label = Combination), hjust = 0, vjust = 0) +
  labs(x = "Sum of Standard Deviations", y = "Mean Score") +
  ggtitle("Top 10 Combinations of 4 Male Athletes with Highest Mean Score")

  # Display the plot
  print(p3)

  # Combine the three plots into one three-panel plot
  combined_plot <- cowplot::plot_grid(p1, p2, p3, ncol = 1)

  # Print the combined plot
  print(combined_plot)
```